---
title: "Practical Machine Learning"
author: "rtaph"
date: "December 16, 2014"
---

******************************
## Title
  
The goal of this project is to predict the manner in which they did the exercise. 
"dataset with 5 classes (sitting-down, standing-up, standing, walking, and sitting) collected on 8 hours of activities of 4 healthy subjects." 


### Data Processing
I begin the analysis by loading libraries and setting a few global parameters:
  
```{r chunkOpts, cache=TRUE}
## load needed libraries, set global options, and working directory
library(knitr); library(plyr)
opts_chunk$set(echo = TRUE)       
setwd("~/Documents/Courses/datasciencecoursera/MachLearnProj/")
```

I then check that the data file exists, download it (if needed), and unzip it:

```{r chuckDownload}
#Download file if it does not exist
if (!file.exists("pml-training.csv")) {
  fileURL <- "http://bit.ly/1GMwAry"
  download.file(fileURL, destfile = "pml-training.csv", method="curl")
}
if (!file.exists("pml-testing.csv")) {
  fileURL <- "http://bit.ly/1znCGyy"
  download.file(fileURL, destfile = "pml-testing.csv", method="curl")
}
```



```{r chunkLoadData, cache=TRUE}
  # Load the data and assign it to a variable
  train    =  read.csv("pml-training.csv", na.strings = "#DIV/0!")
  set.seed(1234); t1   = train[sample(nrow(train),500),]
  test    =  read.csv("pml-testing.csv")
  rbind(training = dim(train), testing = dim(test))
```

With 159 predictor variables available, a model selection method such as best subset selection is unfeasible. 
